### 2014-10-22 paper/include/discussion.tex
disc
wrote & corrections

\textbf{Discuss how 3D DRAM affects everything}
-Will any approach be benefited / dis-benefited by it? 
-Which technique seems future proof?

>>>>

Another proposed technique is ESKIMO by Isen and John \cite{eskimo}, which focus at reducing total DRAM power through optimizing the accesses to the DRAM. By using one bit of a cache row to denote whether it hold nonsense data, it becomes possible to resolve some memory accesses in the caches, which otherwise would have needed to access DRAM. A memory region is regarded as nonsense if it has been deallocated, or allocated but not written to. The ISA has to be extended to provide the allocation and deallocation information. If a block has to be replaced in a write-through cache and the replaced block is dirty but also deallocated, the cache can ignore writing the replaced block to memory. Similarly, a write miss to a newly allocated area results in a read access that can also be ignored. The latter optimization has support of a structure stored in DRAM which tracks nonsense rows.

The 
Combining ESKIMO with the selective refresh implementation proposed by Oshawa \cite{oshawa} (ESKIMO[31]), $39\%$ of total DRAM power was saved on average.

-Will any approach be benefited / dis-benefited by it?
-Which technique seems future proof?

### 2014-10-22 paper/include/secret.tex
wrote & corrections

With SECRET Lin et al. \cite{secret} masks cells with short retention time (bad cells) using Error Correction Pointers (ECP). The bad cells are mapped with a one time profiling done before first system use and all cells that do not meet the retention time requirements gets a ECP. A ECP is basically a pointer to another cell that is healthy. The framework, which is implemented inside the memory controller, checks whether there is any valid ECP for the address is accessed. If there is, it replaces the cell's data with the data from the cell pointed to by the ECP. The ECPs are stored in the DRAM during runtime and a cache in the memory controller is used to mask the delay of fetching the ECPs from DRAM.

The framework divides the DRAM into regions of typically \textit{128 KB}. For every one of these regions a \textit{Directory Entry} is kept. These directories holds the number of bad cells in the region and a address to where the ECPs are stored. As the Directory Entries need to be stored in a contiguous address space and the probability of that existing without any bad cell three copies are stored. A majority vote is then performed to mask any bit error in the Directory Entry. The size of each entry is $36\:b$ and all entries in a \textit{4 GB} system thus generates a total of $3 \times 36\:b \times \frac{4\:GB}{128\:KB} =$ \textit{432 KB} $= 0.010\%$ storage overhead. This is without the space needed for the actual ECPs which requires $21\:b$ each and the total space thereby varies with the number of bad cells.

This approach gives a high energy saving of up to $87.2\%$ in refresh power and a reduction of $18.57\%$ in total DRAM power consumption with the ECP cache overhead taken into account and the refresh rate set to $512\:ms$ instead of the default $64\:ms$. The performance impact ranges from $1.3\%$ degradation at the worst to a $1.4\%$ improvement at the best. The degradation of performance is due to the added memory access needed to fetch the ECPs, and the performance improvement comes from the decreased congestion due to less refresh operations performed.

>>>>

With SECRET Lin et al. \cite{secret} masks cells with short retention time using Error Correction Pointers (ECP). The DRAM cells are mapped through profiling, which is performed the first time the system is in use, and all cells that do not meet the retention time requirements is assigned an ECP. An ECP is basically a pointer to a cell that met the retention time requirements. The framework logic, which is implemented inside the MC, checks whether there is any ECP for the address accessed. If there is, it replaces the cell's data with the data from the cell pointed to by the ECP. The ECPs are stored in DRAM during runtime and are also cached in the MC to mask the delay of fetching them from DRAM.

The framework divides the DRAM into regions of typically \textit{128 KB}. For every region a \textit{Directory Entry} is kept. These directories holds the number of bad cells in a region and an address to where the ECPs are stored. As the Directory Entries need to be stored in a contiguous address space and the probability of that existing without any bad cell, three copies are stored. A majority vote is then performed to mask any bit error in the Directory Entries. The size of each entry is $36\:b$ and all entries in a \textit{4 GB} system thus generates a total of $3 \times 36\:b \times \frac{4\:GB}{128\:KB} =$ \textit{432 KB} $= 0.010\%$ storage overhead. This is without the space needed for the actual ECPs which requires $21\:b$ each and the total space thereby varies with the number of bad cells.

This technique saves up to $87.2\%$ of refresh power consumption which corresponds to a total reduction of DRAM power consumption of $18.6\%$, when the ECP cache overhead is taken into account and the refresh rate set to $512\:ms$ instead of the default $64\:ms$. The performance impact ranges from $1.3\%$ degradation at the worst to a $1.4\%$ improvement at the best. The degradation is due to the added memory accesses needed to fetch the ECPs, and the  improvement comes from the decreased DRAM access congestion due to less refresh operations performed.

### 2014-10-22 paper/include/smart_refresh.tex
wrote & corrections

Smart Refresh modifies the MC but does not change the DRAM module nor the interface between them. The modifications of the MC constitute of counter update logic, the counter array, and a pending refresh request queue. The memory footprint of the modifications is $0.0048\%$ of total device capacity when three-bits counters are used assuming a cache row size of \textit{8~KB}.

>>>>

Smart Refresh modifies the MC but does not change the DRAM module nor the interface between them. The modifications of the MC constitute of counter update logic, the counter array, and a pending refresh request queue. The memory footprint of the modifications is $0.0048\%$ of total device capacity when three-bits counters are used assuming a DRAM row size of \textit{8~KB}.
